MULTI EPOCH SPATIALLY RESOLVED RADIO OBSERVATIONS OF BETELGEUSE'S WIND ACCELERATION REGION

### 2014-06-10 draft1/ms.tex
3rd commit
First from Onsala

\section{INTRODUCTION}
The problem with early M supergiants and introduce Betelgeuse

The wind acceleration region importance and thermal free-free as a probe

Previous radio observations (earlier VLA in agreement with ALfven models, skinner, lim, richards)

@@

\begin{deluxetable*}{lcccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{Multi Epoch VLA A-Configuration Plus Pie Town Link Observations of Betelgeuse.}
\tablehead{\colhead{Date}						&
			\colhead{Wavelength}			    &
           	\colhead{Restoring Beam}            &
           	\colhead{Image rms}            		&
           	\colhead{$\theta _{\mathrm{maj}}$}  &
			\colhead{$\theta _{\mathrm{maj}}/\theta _{\mathrm{min}}$}	&
           	\colhead{P.A.}  &
          	\colhead{$F_{\nu}$}  &
           	\colhead{$T_{b}$}  \\
	\colhead{}		                			& 
	\colhead{(cm)}                         		& 
	\colhead{(mas $\times$ mas)}   	&
	\colhead{(mJy/Beam)}                    	&
	\colhead{(mas)}   		    &
	\colhead{}   		    &
	\colhead{(deg)}   		    &
	\colhead{(mJy)}   		    &
	\colhead{(K)}		}
\startdata
		 2004 Oct 21,30  & 0.7 & $39\times 26$ & 0.37	& $99\pm 3$ & $0.93\pm 0.04$& $92\pm 20$ &$28.67\pm 0.53$ & \\
		 				 & 1.3 & $80\times 42$ & 0.09   & $121\pm 2$& $\dots$ & $\dots$& $13.88\pm 0.10$& \\
						 & 2.0 & $121\times 91$ & 0.08  & $158\pm 6$& $\dots$ & $\dots$ & $7.23\pm 0.15$& \\
						 & 3.6 & $208\times 126$ & 0.02	& $215\pm 7$& $0.87\pm 0.04$ & $162\pm 7$& $3.34\pm 0.03$& \\
						 & 6.1 & $377\times 264$ & 0.02	& $315\pm 30$& $0.59\pm 0.13$ & $173\pm 10$ & $1.55\pm 0.04$&\\
						 & 20.5& $1262\times 889$ & 0.03& $\le 889$ & $\dots$ & $\dots$ &$0.25\pm 0.03$ &\\
\hline
\rule{-2.6pt}{2.5ex}  2003 Aug 10,12 & 0.7 		& &	& $103\pm 4$& $0.89\pm 0.06$ & $104\pm 16$& $28.05\pm 0.84$& \\
									 & 1.3		& &	& $122\pm 5$& $\dots$& $\dots$& $11.20\pm 0.24$& \\
									 & 2.0		& &	& $132\pm 10$& $0.87\pm 0.10$& $11\pm 27$& $5.88\pm 0.17$&\\
									 & 3.6		& &	& $193\pm 7$& $0.73\pm 0.06$ & $152\pm 7$& $2.80\pm 0.04$& \\
									 & 6.1 		& &	& $209\pm 49$& $\dots$ & $\dots$& $1.22\pm 0.04$& \\
					                 &20.5		& &	& $\dots$& $\dots$ & $\dots$& $0.26\pm 0.03$& \\
\hline
\rule{-2.6pt}{2.5ex}  2002 Apr 12,13  & 1.3 		& &	&$134\pm 9$ & $76 \pm 7$& $36\pm 10$& $8.96\pm 0.24$& \\
							& 2.0		& &	& $166\pm 16$& $63\pm 10$&$41\pm 11$ & $5.32\pm 0.23$& \\
							& 3.5		& &	& $234\pm 9$& $73\pm 5$& $40\pm7$& $2.66\pm 0.04$& \\
							& 6.1 	& &	& $\dots$ & $\dots$& $\dots$& $\dots$& $\dots$\\
							& 20.5	& &	& $\dots$ & $\dots$ & $\dots$& $0.38\pm 0.06$& \\
\hline
\rule{-2.6pt}{2.5ex}  2002 Feb 17,18 & 1.3 		& &	& & & & &\\
									& 2.0		& &	& & & & & \\
									& 3.6		& &	& & & & & \\
									& 6.1 		& &	& & & & & \\
									& 20.5		& &	& & & & & \\
\hline
\rule{-2.6pt}{2.5ex}  2001 Jan 02  & 1.3 		& &	& & & & & \\
 		 2000 Dec 23 & 0.7		& &	& & & & & \\
\hline
\rule{-2.6pt}{2.5ex}  1998 Mar 29,30 & 0.7 		& &	& & & & & \\
									 & 1.3		& & 	&  & & & &
\enddata
\tablenotetext{}{Notes.- Restoring beam size obtained using uniform weighting. The P.A. is measured in degrees east of north. }
\label{tab:tab1}
\end{deluxetable*}

@@

\section{RESULTS} 
\subsection{Visibilities} 
\subsection{Radio Maps}

>>>>

\section{INTRODUCTION}
Red supergiants (RSGs) lose mass to the interstellar medium in the form of a massive ($\dot{M} \sim 10^{-7} - 10^{-4}\,M_{\odot}\,\rm{yr}^{-1}$) cool wind, with terminal velocities ($10\lesssim v_{\infty}\lesssim 50\,{\rm{km}}\,\rm{s}^{-1}$) typically less than the photospheric escape velocity  ($v_{\rm{esc}} \sim 100\,\rm{km},\rm{s}^{-1}$). These winds are major contributors of heavy elements to the interstellar medium and play a crucial role in stellar evolution \citep{chiosi_1986}, and also in explaining the frequency of supernovae in the galaxy \citep[e.g.,][]{van_loon_2010}. Despite their importance, the mechanisms responsible for the formation of these winds in late K and early M-type RSGs remain largely unknown. Dust is observed too far from the star to play a significant role in the mass-loss process \citep{danchi_1994} and pulsation amplitudes are too low to initiate the mass-loss \citep{smith_1989}. Magnetohydrodynamic (MHD) waves \citep[e.g.,][]{thirumalai_2012} and large convective cells \citep[e.g.,][]{josselin_2007} have been proposed as alternative potential mass-loss mechanisms in RSGs but spatially resolved multi-epoch observations of RSGs are required to test these competing theories.

Betelgeuse ($\alpha$ Ori, M1-2Ia-Ibe) is the closest isolated RSG at a distance of $197 \pm 45$\,pc \citep{harper_2008} and is therefore the prototype of the late K and early M RSG class. Its large angular diameter ($\phi _{\star} = 42.49 \pm 0.06\,$mas in the K band, \citealt{ohnaka_2011}) coupled with its extended atmosphere makes its an excellent target for detailed multi-wavelength studies aiming to build a complete understanding of mass-loss in RSGs. As a result it has recently been the focus of multiple high spatial resolution studies tracing various spatial scales.

Two distinct flows in its circumstellar envelope (CSE) were imaged at 1.3\,mm by \cite{ogorman_2012} which traced CO($J=2-1$) on scales between $\sim 40$\,R$_{\star}$ and $\sim 750$\,R$_{\star}$. These observations revealed an irregular CSE with a notable asymmetry in the S-W direction extending out to $\sim 200$\,R$_{\star}$. Thermal infrared images using the VLT/VISIR ($\lambda= 8 - 20\,\mu$m, \citealt{kervella_2011} detected an envelope of inhomogeneous surface brightness up to 100\,R$_{\star}$, whose spectral energy distribution is typical of oxygen-rich dust. These images also uncovered a ring-like structure at radius $20-45$\,R$_{\star}$ which may be related to the dust condensation radius. The inner CSE  was probed in the near-IR with the VLT/NACO by \cite{kervella_2009} who discovered a molecular plume extending out to almost 6\,R$_{\star}$, which they attributed to the action of a giant convection cell. Indeed Haubois et al.'s (2009) image reconstructions indicate two bright spots on the surface of Betelgeuse which may be a signature of such convective cells.

Wind acceleration zone. The wind acceleration region importance and thermal free-free as a probe. Previous radio observations (earlier VLA in agreement with ALfven models, skinner, lim, richards)

@@

\begin{deluxetable*}{lcccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{Multi Epoch VLA A-Configuration Plus Pie Town Link Observations of Betelgeuse.}
\tablehead{\colhead{Date}						&
			\colhead{Wavelength}			    &
           	\colhead{Restoring Beam}            &
           	\colhead{Image rms}            		&
           	\colhead{$\theta _{\mathrm{maj}}$}  &
			\colhead{$\theta _{\mathrm{maj}}/\theta _{\mathrm{min}}$}	&
           	\colhead{P.A.}  &
          	\colhead{$F_{\nu}$}  &
           	\colhead{$T_{b}$}  \\
	\colhead{}		                			& 
	\colhead{(cm)}                         		& 
	\colhead{(mas $\times$ mas)}   	&
	\colhead{(mJy/Beam)}                    	&
	\colhead{(mas)}   		    &
	\colhead{}   		    &
	\colhead{(deg)}   		    &
	\colhead{(mJy)}   		    &
	\colhead{(K)}		}
\startdata
		 2004 Oct 21,30  & 0.7 & $39\times 26$ & 0.37	& $99\pm 3$ & $0.93\pm 0.04$& $92\pm 20$ &$28.67\pm 0.53$ & $2940\pm 170$\\
		 				 & 1.3 & $80\times 42$ & 0.09   & $121\pm 2$& $\dots$ & $\dots$& $13.88\pm 0.10$& $3140\pm 80$\\
						 & 2.0 & $121\times 91$ & 0.08  & $158\pm 6$& $\dots$ & $\dots$ & $7.23\pm 0.15$& $2270\pm 130$ \\
						 & 3.6 & $208\times 126$ & 0.02	& $215\pm 7$& $0.87\pm 0.04$ & $162\pm 7$& $3.34\pm 0.03$& $2110\pm 110$\\
						 & 6.1 & $377\times 264$ & 0.02	& $315\pm 30$& $0.59\pm 0.13$ & $173\pm 10$ & $1.55\pm 0.04$& $1140\pm 160$\\
						 & 20.5& $1262\times 889$ & 0.03& $\le 889$ & $\dots$ & $\dots$ &$0.25\pm 0.03$ &$ \textgreater 260$\\
\hline
\rule{-2.6pt}{2.5ex}  2003 Aug 10,12 & 0.7 		& $40\times 27$ & 0.46	& $103\pm 4$& $0.89\pm 0.06$ & $104\pm 16$& $28.05\pm 0.84$& $2760\pm 230$ \\
									 & 1.3		& $80\times 42$ & 0.17	& $122\pm 5$& $\dots$& $\dots$& $11.20\pm 0.24$& $2490\pm 150$\\
									 & 2.0		& $119\times 96$& 0.10	& $132\pm 10$& $0.87\pm 0.10$& $11\pm 27$& $5.88\pm 0.17$&$3040\pm 360$\\
									 & 3.6		& $ 204\times 139$&	0.03& $193\pm 7$& $0.73\pm 0.06$ & $152\pm 7$& $2.80\pm 0.04$&  $2610\pm 170$\\
									 & 6.1 		& $378\times 297$&	0.03& $209\pm 49$& $\dots$ & $\dots$& $1.22\pm 0.04$&$2040\pm 680$ \\
					                 &20.5		& $1247\times 931$&0.04	& $\le 931$& $\dots$ & $\dots$& $0.26\pm 0.03$& $ \textgreater 250$\\
\hline
\rule{-2.6pt}{2.5ex}  2002 Apr 12,13  & 1.3 		& $91\times 59$ & 0.18	&$134\pm 9$ & $0.76 \pm 0.07$& $36\pm 10$& $8.96\pm 0.24$& $2170\pm 250$\\
							& 2.0		&$131\times 98$ & 0.39	& $166\pm 16$& $0.63\pm 0.10$&$41\pm 11$ & $5.32\pm 0.23$& $2420\pm 450$ \\
							& 3.6		& $224\times 155$&0.03	& $234\pm 9$& $0.73\pm 0.05$& $40\pm7$& $2.66\pm 0.04$& $1690\pm 110$\\
							& 20.5	& $1398\times 1146$& 0.06	& $\le 1146$ & $\dots$ & $\dots$& $0.38\pm 0.06$& $ \textgreater 240$\\
\hline
\rule{-2.6pt}{2.5ex}  2002 Feb 17,18 & 1.3 		& $83\times 48$&0.14	&$120\pm 4$ & $0.91\pm 0.04$& $30\pm 13$&$10.87\pm 0.17$ & $2750\pm 140$\\
									& 2.0		& $128\times 90$&0.11	&$140\pm 13$& $\dots$& $\dots$&$5.38\pm 0.22$ & $2150\pm 300$\\
									& 3.6		& $200\times 135$&	0.03&$199\pm 8$ & $\dots$& $\dots$ & $2.85\pm 0.04$& $1830\pm 110$\\
									& 6.1 		& &	& $\dots$& $\dots$& $\dots$& $\dots$& $\dots$\\
									& 20.5		&$1312\times 951$ &0.05	& $\le 951$& $\dots$& $\dots$& $0.30\pm 0.05$& $ \textgreater 270$\\
\hline
\rule{-2.6pt}{2.5ex}  2001 Jan 02  & 1.3 		& $78\times 42$&0.08	&$124\pm 2$ & $0.92\pm 0.02$& $40\pm 8$& $12.58\pm0.08$& $2920\pm 70$ \\
 		 2000 Dec 23 & 0.7		& $44\times 20$& 0.18	&$98\pm 2$ &$0.92\pm 0.02$ &$0\pm 7$ & $29.02\pm 0.30$& $3070\pm 100$
%\hline
%\rule{-2.6pt}{2.5ex}  1998 Mar 29,30 & 0.7 		& &	& $104\pm 6$& $0.88\pm 0.07$& $118\pm 20$& $18.19\pm 0.73$& $1770\pm 180$ \\
%									 & 1.3		& & 	& $157\pm 8$ & $0.61\pm 0.06$& $46\pm 6$&$10.99\pm 0.31$ & $2420\pm 250$
\enddata
\tablenotetext{}{Notes.- The restoring beam and image rms noise values were obtained using uniform weighting and include the Pie Town antenna baselines. The position angles (measured in degrees east of north) of the restoring beams in these images are not given here but are all between $30^{\circ}$ and $70^{\circ}$. The major axis of the stellar radio disk, $\theta _{\mathrm{maj}}$, the axis ratio of the major and minor radio disks, $\theta _{\mathrm{maj}}/\theta _{\mathrm{min}}$, the position angle, P.A., and the total flux density, $F_{\nu}$, are all derived from the best-fit uniform-brightness ($T_b$) elliptical-disk models.}
\label{tab:tab1}
\end{deluxetable*}

@@

\section{RESULTS} 
\subsection{Visibilities} 
\subsection{Radio Images} 
discuss fast switching (carilli paper)
\subsection{Wind Thermal Profile}
give formula
